---
title: "Course_Project_1"
author: "suveyda"
date: "January 20, 2017"
output: html_document
---

# Course Project 1

## Exploratory Data Analysis by Johns Hopkins University

### Introduction

In this course project our overall goal is to examine how household energy usage varies over a 2-day period in February, 2007. Our task is to reconstruct the given plots using the base plotting system.

### Data

This assignment uses data from the UC Irvine Machine Learning Repository, a popular repository for machine learning datasets. In particular, we will be using the "Individual household electric power consumption Data Set" which is available on the course web site.

The given dataset contains measurements of electric power consumption in one household with a one-minute sampling rate over a period of almost 4 years. Different electrical quantities and some sub-metering values are available.
The following descriptions of the 9 variables in the dataset are taken from the UCI web site:

* Date: Date in format dd/mm/yyyy
* Time: time in format hh:mm:ss
* Global_active_power: household global minute-averaged active power (in kilowatt)
* Global_reactive_power: household global minute-averaged reactive power (in kilowatt)
* Voltage: minute-averaged voltage (in volt)
* Global_intensity: household global minute-averaged current intensity (in ampere)
* Sub_metering_1: energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered).
* Sub_metering_2: energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator and a light.
* Sub_metering_3: energy sub-metering No. 3 (in watt-hour of active energy). It corresponds to an electric water-heater and an air-conditioner.

Let's first load the dataset into R:

```{r}
data <- read.table("household_power_consumption.txt", header = TRUE, sep = ";", stringsAsFactors=FALSE)
```

As recommended in the project, we convert the Date and Time variables to Date/Time classes in R using the strptime()  and as.Date() functions. Furthermore, since we will only be using data from the dates 2007-02-01 and 2007-02-02, we create a subset of the data below.

```{r}
data$Date_Time <- paste(data$Date, data$Time)
data$Date_Time <- strptime(data$Date_Time, "%d/%m/%Y %H:%M:%S")

data$Date <- as.Date(data$Date, "%d/%m/%Y")
days <- as.Date(c("2007-02-01", "2007-02-02"))
sub.data <- subset(data, Date >= days[1] & Date <= days[2])
```

### Plot 1

```{r}
#png("plot1.png", width=480, height=480)
hist(as.numeric(sub.data$Global_active_power), 
	main = "Global Active Power", 
	xlab = "Global Active Power (kilowatts)",
	ylim = c(0,1200), col = "red")
#dev.off()
```

### Plot 2

```{r}
#png("plot2.png", width=480, height=480)
plot(sub.data$Date_Time, sub.data$Global_active_power, type = "l",
	xlab = "", ylab = "Global Active Power (kilowatts)")
#dev.off()
```

### Plot 3

```{r}
#png("plot3.png", width=480, height=480)
plot(sub.data$Date_Time, as.numeric(sub.data$Sub_metering_1), 
	xlab = "", ylab = "Energy sub metering", type = "n")
lines(sub.data$Date_Time, as.numeric(sub.data$Sub_metering_1), col = "black")
lines(sub.data$Date_Time, as.numeric(sub.data$Sub_metering_2), col = "red")
lines(sub.data$Date_Time, sub.data$Sub_metering_3, col = "blue")
legend("topright", lty = 1, col = c("black", "red", "blue"), 
	legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"))
#dev.off()
```

### Plot 4

```{r}
#png("plot4.png", width=480, height=480)
par(mfrow=c(2,2))

with(sub.data, {
	plot(Date_Time, Global_active_power, type = "l", 
	  xlab = "", ylab = "Global Active Power")
	
	plot(Date_Time, Voltage, type = "l", xlab = "datetime", ylab = "Voltage")

	plot(sub.data$Date_Time, as.numeric(sub.data$Sub_metering_1), 
		xlab = "", ylab = "Energy sub metering", type = "n")
	lines(sub.data$Date_Time, as.numeric(sub.data$Sub_metering_1), col = "black")
	lines(sub.data$Date_Time, as.numeric(sub.data$Sub_metering_2), col = "red")
	lines(sub.data$Date_Time, sub.data$Sub_metering_3, col = "blue")
	legend("topright", lty = 1, cex = 0.8, col = c("black", "red", "blue"),
		legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"))

	plot(Date_Time, Global_reactive_power,  type = "l",
	     xlab = "datetime", ylab="Global_reactive_power")
})
#dev.off()
```
